What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-06 android-stacks.tex
Initial commit

\begin{itemize}
\item  A multitude undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform).
\item  Odd exception chains (e.g., Error wrapping checked and runtime exceptions and vice-versa). As an indication that the purpose of Java hybrid exception model may not have been adequately used.
\item  Stack traces caused by uncaught checked exceptions.IMPROVE.

\end{itemize}

@@

\textbf{5) Document all exceptions thrown by a method}I~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}). The exceptions thrown by a method are an important part of methods interface, and is required to use the method properly. The checked exceptions are already part of the  methods signature, and the method caller is aware of the checked exceptions being thrown by it. According to~\cite{bloch2008effective}, it is also wise to document the explicitly signaled runtime exceptions\footnote{This excludes the implicitly signaled JVM runtime exceptions related to coding errors} as carefully as checked exceptions. Doing so, the clients of a method will be aware of the exceptions the method can throw. If the developer fails to do follow this practice (specially when developing library code) it will be difficult or even impossible for the caller to make effective use of such method ~\cite{wirfs2006toward, bloch2008effective}.

@@

\noindent\emph{Exception Signaler Classifier} After its initialization (see Section 
set of packages for each classification category and pattern matches the signaler name
to classify it to one of the categories.

@@

To answer this question, we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler. The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.~\todo{Move this to the previous section? IMO, It is better to keep all facts about how we did the
research in one place}

@@

As detailed in Section III, 443 Android projects were carefully selected to enable a deaper investigation of stack trace information, by leveraging stack trace information with the source code and bytecode analysis. For this subset of Android projects we could extract every package related to each exception signaler, and based on this information we could classify every stack trace based on its root signaler as presented in Table~\ref{signalers}.

@@

For each exception we could discover: the number of times it occurred on stacks, the number of distinct projects such stacks were defined (and consequently its popularity) and the of times it was signaled by different signaler types. We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (443). Table~\ref{toptenandroid} and Figuere~\ref{androidsignaler} presents the mined data.

@@

\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset. For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification].}

@@

TODO: Include a Histogram here.

@@

To answer this question we analyzed general characteristics of stack traces reported on Java repositories (e.g., exceptions being thrown presented in Tables~\ref{toptenjava} and ~\ref{toptenandroid}), and took a deeper look at the stack traces of the Android repositories combining them with with source code and byte code analyses. Doing so, we mined some characteristics of stacks that points to best practices violations.

@@

From Tables Tables~\ref{toptenjava}, ~\ref{toptenandroid} we can observe that, in both Java and Android related issues, direct instances java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and 18,88\% of Android repositories respectively). Moreover, after inspecting the top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that stack traces on which direct instances of java.lang.Exception and java.lang.Error were thrown. Throwing general exceptions is usually considered a bad practice, because the type does not carry enough information to identify the cause of the exceptional behaviour. In such cases, developers need to relying on exception message which can be neither complete nor precise ~\cite{gosling2000java}.

@@

\subsection{Undocumented runtime exceptions signaled by third party code}

@@

As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions. 
When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code. Only manual code inspection can reveal the reason. However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented. Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code. As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.

On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions. 
It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients. Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Exception Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
    Runtime	&	None	&	2360	&	359	\\
Runtime	&	Runtime	&	560	&	182	\\
Checked	&	None	&	422	&	117	\\
Error	&	None	&	381	&	141	\\
Runtime	&	Checked	&	109	&	70	\\
Checked	&	Checked	&	98	&	42	\\
Runtime	&	Error	&	55	&	38	\\
Checked	&	Runtime	&	22	&	12	\\
Error	&	Error	&	15	&	14	\\
Undefined	&	None	&	15	&	10	\\
Error	&	Runtime	&	13	&	6	\\
Checked	&	Error	&	8	&	7	\\
Error	&	Checked	&	7	&	4	\\
Throwable	&	Runtime	&	3	&	1	\\
Undefined	&	Runtime	&	3	&	3	\\
Throwable	&	None	&	1	&	1	\\
Undefined	&	Error	&	1	&	1	\\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 443 ANDROID projects found on Github}
\label{fig:wrappingandroid}
\end{table}

@@

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12} investigated the degree of reuse across applications in Android Market, the study showed that almost 23\% of the classes inherited
from a base class in the Android API, and that 217 mobile apps were reused
completely by another mobile app. Pathak et al.~\cite{Patha11} analyzed bug
reports and developers discussions of Android platform and found out that
approximately 20\% of energy-related bugs in Android occurred after an OS
update. McDonnell et al.~\cite{McDon13} conducted a case study of the
co-evolution behavior of Android API and 10 dependent applications using the
version history data found in github. The study found that approximately 25\% of
all methods in the client code used the Android API, and that the methods
reusing fast-evolving APIs were more defect prone then others. Vásquez et
al.~\cite{Linar13} analyzed approximately 7K free Android apps and observed that
the last successful apps used Android APIs that were on average 300\% more
change-prone than the APIs used by the most successful apps. Pingyu and
Elbaum~\cite{Zhang12} analyzed bug reports of 5 Android applications an observed
that 29\% had to do with poor exceptional handling code.  Our work differs from
the others as it aims at distilling the information of bug reports describing
uncaught exceptions created for mobile applications in Github and Googlecode, in
order to get a first view of what is causing the crashes across applications
available and what the characteristics of the stack traces can tell us about the
exception structure of those applications. Moreover this work\ldots.

>>>>

\begin{itemize}
\item  A multitude undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform).
\item  Odd exception chains (e.g., Error wrapping checked and runtime exceptions and vice-versa). As an indication that the purpose of Java hybrid exception model may not have been adequately used.
\item  Stack traces caused by uncaught checked exceptions.
\end{itemize}

@@

\textbf{5) Document all exceptions thrown by a method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}). The exceptions thrown by a method are an important part of methods interface, and is required to use the method properly. The checked exceptions are already part of the  methods signature, and the method caller is aware of the checked exceptions being thrown by it. According to~\cite{bloch2008effective}, it is also wise to document the explicitly signaled runtime exceptions\footnote{This excludes the implicitly signaled JVM runtime exceptions related to coding errors} as carefully as checked exceptions. Doing so, the clients of a method will be aware of the exceptions the method can throw. If the developer fails to do follow this practice (specially when developing library code) it will be difficult or even impossible for the caller to make effective use of such method ~\cite{wirfs2006toward, bloch2008effective}.

@@

\noindent\emph{Exception Signaler Classifier} After its initialization it classifies each signaler according to its origin (i.e. Android Plaform, Aplication, Library, Libcore, or Java) based on pattern matching between the signaler name
and the packages associated to each category.

@@

To answer this question, we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler. The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.
%~\todo{Move this to the previous section? IMO, It is better to keep all facts about how we did the research in one place}

@@

As detailed in Section III, 482 Android projects were carefully selected to enable a deaper investigation of stack trace information, by leveraging stack trace information with the source code and bytecode analysis. For this subset of Android projects we could extract every package related to each exception signaler, and based on this information we could classify every stack trace based on its root signaler as presented in Table~\ref{signalers}.

@@

For each exception we could discover: the number of times it occurred on stacks, the number of distinct projects such stacks were defined (and consequently its popularity) and the of times it was signaled by different signaler types. We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (482). Table~\ref{toptenandroid} and Figuere~\ref{androidsignaler} presents the mined data.

@@

%commented only for submission:
%\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset. For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification].}

@@

%TODO: Include a Histogram here.

@@

To answer this question we analyzed general characteristics of stack traces reported on Java repositories (e.g., exceptions being thrown presented in Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid}), and took a deeper look at the stack traces of the Android repositories combining them with with source code and byte code analyses. Doing so, we mined some characteristics of stacks that points to best practices violations.

@@

From Tables~\ref{tab:toptenjava}, ~\ref{tab:toptenandroid} we can observe that, in both Java and Android related issues, direct instances java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and 18,88\% of Android repositories respectively). Moreover, after inspecting the top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that stack traces on which direct instances of java.lang.Exception and java.lang.Error were thrown. Throwing general exceptions is usually considered a bad practice, because the type does not carry enough information to identify the cause of the exceptional behaviour. In such cases, developers need to relying on exception message which can be neither complete nor precise ~\cite{gosling2000java}.

@@

%todo{DISCOMMENT AFTER SUBMISSION}
%\subsection{Undocumented runtime exceptions signaled by third party code}

%In this study we also classified each stack trace according to the type of the root exception been signaled. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on the Android repositories.

%todo{discomment after submission}
%As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions. When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code. Only manual code inspection can reveal the reason. However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented. Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code. As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.

%On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions. It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients. Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.

@@

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Exception Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
    Runtime	&	None	&	2360	&	359	\\
Runtime	&	Runtime	&	560	&	182	\\
Checked	&	None	&	422	&	117	\\
Error	&	None	&	381	&	141	\\
Runtime	&	Checked	&	109	&	70	\\
Checked	&	Checked	&	98	&	42	\\
Runtime	&	Error	&	55	&	38	\\
Checked	&	Runtime	&	22	&	12	\\
Error	&	Error	&	15	&	14	\\
Undefined	&	None	&	15	&	10	\\
Error	&	Runtime	&	13	&	6	\\
Checked	&	Error	&	8	&	7	\\
Error	&	Checked	&	7	&	4	\\
Throwable	&	Runtime	&	3	&	1	\\
Undefined	&	Runtime	&	3	&	3	\\
Throwable	&	None	&	1	&	1	\\
Undefined	&	Error	&	1	&	1	\\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
\label{fig:wrappingandroid}
\end{table}

@@

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12} investigated the degree of reuse across applications in Android Market, the study showed that almost 23\% of the classes inherited
from a base class in the Android API, and that 217 mobile apps were reused
completely by another mobile app. Pathak et al.~\cite{Patha11} analyzed bug
reports and developers discussions of Android platform and found out that
approximately 20\% of energy-related bugs in Android occurred after an OS
update. McDonnell et al.~\cite{McDon13} conducted a case study of the
co-evolution behavior of Android API and 10 dependent applications using the
version history data found in github. The study found that approximately 25\% of
all methods in the client code used the Android API, and that the methods
reusing fast-evolving APIs were more defect prone then others. Vásquez et
al.~\cite{Linar13} analyzed approximately 7K free Android apps and observed that
the last successful apps used Android APIs that were on average 300\% more
change-prone than the APIs used by the most successful apps. Pingyu and
Elbaum~\cite{Zhang12} analyzed bug reports of 5 Android applications an observed
that 29\% had to do with poor exceptional handling code.  Our work differs from
the others as it aims at distilling the information of bug reports describing
uncaught exceptions created for mobile applications in Github and Googlecode, in
order to get a first view of what is causing the crashes across applications
available and what the characteristics of the stack traces can tell us about the
exception structure of those applications. %Moreover this work\ldots.
